---
title: "nitrateApp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nitrateApp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width = 6, fig.height = 3
)
```

## Overview
Rising river nitrate concentrations are becoming increasingly threatening to aquatic ecosystems. The nitrateApp package provides a clean dataset and interactive app to explore and visualise nitrate concentrations for three NEON sites between June - July 2019. This vignette demonstrates how to:

- Load and inspect the dataset
- Use simple functions to explore the data
- Run the interactive Shiny app

## Loading the package and data 
```{r}
library(nitrateApp)
data("nitrates_small", package = "nitrateApp")
str(nitrates_small)
```

The `nitrates_small` dataset is a small version of nitrate concentration datasets available from NEON. These sites differ in nitrate concentrations due to their individual geographic and aquatic characteristics.
These differences and other patterns can be further explored using the `nitrates_small` dataset. 

## Helper examples to explore the dataset

Calculating some summary statistics
```{r, message=FALSE, warning=FALSE}
library(dplyr)

nitrates_small |>
  group_by(site)|>
  summarise(
    mean_mgL = mean(nitrate_mgL, na.rm = TRUE),
    min_mgL = min(nitrate_mgL, na.rm = TRUE),
    max_mgL = max(nitrate_mgL, na.rm = TRUE)
  )
```

## Visualising the data
```{r, message=FALSE, warning=FALSE}
library(ggplot2)

ggplot(nitrates_small, aes(date, nitrate_mgL, color = site)) +
geom_line() +
geom_point() +
labs(x = NULL, y = "Nitrate (mg/L)", title = "Daily nitrate concentrations by site") +
theme_minimal()
```

## Launch the interactive app 

To explore the data interactively run: 
```{r, eval=FALSE}
nitrateApp::run_nitrate_app()
```

With the app you can: 

- Select one or more site(s)
- Select a date range
- View an updating time series for the selected date range and location
